---
title: "Fractals with R"
author: "R. G. Cronce"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Fractals with R

A five part series found somewhere on the web.  Found these when I was looking for code
to generate a Koch snowflake and found the others along with it.

[Part 1: Introduction](https://bmscblog.wordpress.com/2012/08/22/fractal-graphics-with-r/)

[Part 2: The Sierpinski Carpet](https://bmscblog.wordpress.com/2012/10/29/fractal-graphics-with-r-part-2/)

[Part 3: The Hilbert Curve](https://bmscblog.wordpress.com/2012/12/17/fractals-with-r-part-3-the-hilbert-curve/)

[Part 4: The Koch Snowflake](https://bmscblog.wordpress.com/2013/10/04/fractals-with-r-part-4-the-koch-snowflake/)

[Part 5: The Sierpinski Carpet with ggplot2](https://bmscblog.wordpress.com/2014/03/12/fractals-with-r-part-5-sierpinski-carpet-with-ggplot2/)


## Part 1: Introduction

```{r}
IterateTriangle <- function(A){
    B <- cbind(A,0*A);
    C <- cbind(A,A);
    D <- rbind(B,C);
return(D);
}

par(mfrow=c(2,2))
for (i in 1:4){
    T <- matrix(1,1,1)
    for (i in 1:i) T <- IterateTriangle(T);
    image(T,col=c("white","black"),axes=FALSE);
    text(0,1,i,col="black") 
}

```

## Part 2: The Sierpinski Carpet

```{r}
IterateCarpet <- function(A){
B <- cbind(A,A,A);
C <- cbind(A,0*A,A);
D <- rbind(B,C,B);
return(D);
}

S <- matrix(1,1,1);
for (i in 1:5) S <- IterateCarpet(S);

image(S,col=c(0, 12), axes=FALSE)

```

## Part 3: The Hilbert Curve


```{r}
#Written by Allan Roberts, 2012.
hilbert.curve <- function(n){

Double <- function(A){
#The matrix for “H(n)” is equal to “Double(H(n-1))”.

m <- dim(A)[1];
n <- dim(A)[2];
N <- m*n;
B <- A+N;
C <- B+N;
D <- C+N;
E <- cbind(rbind(B,skew.transpose(A)),rbind(C,t(D)));
return(E);
}

Rotate <- function(A){
#Rotates the matrix A clockwise.

m <- dim(A)[1];
n <- dim(A)[2];
N <- m*n;
B <- matrix(0,m,n);
for (i in 1:m) for (j in 1:n) B[j,n+1-i] <- A[i,j]
return(B);
}

skew.transpose <- function(A){
return(Rotate(Rotate(t(A))));
}

rowofx <- function(A,x){
  
#Returns the row index of the matrix A for entry equal to x.
m <- dim(A)[1];
n <- dim(A)[2];
for (i in 1:m) for (j in 1:n) if (A[i,j]==x) return(i);
}

colofx <- function(A,x){

#Returns the column index of the matrix A for entry equal to x.
m <- dim(A)[1];
n <- dim(A)[2];
for (i in 1:m) for (j in 1:n) if (A[i,j]==x) return(j);
}

Draw <- function(A){
#Draws a graphical representation of the matrix A.
A <- Rotate(A);
m <- dim(A)[1];
n <- dim(A)[2];
N <- m*n;
plot(  (rowofx(A,1)-1)/n, (colofx(A,1)-1)/n, pch=19,cex=0.5,ylim = c(0,1), xlim =c(0,1),     ylab=character(1),xlab=character(1),axes=FALSE);
d <- 1/n;
for (i in 1:(N-1)) lines(c((rowofx(A,i)-1)/n,((rowofx(A,i+1)-1)/n)), c((colofx(A,i)-1)/n,((colofx(A,i+1)-1)/n)),lwd=1);
points((rowofx(A,N)-1)/n, (colofx(A,N)-1)/n, pch=19,cex=0.5);
}

H <- function(n){
#H(1) is shown in Figure 2.
if (n==0) return(matrix(c(2,1,3,4),2,2));
return(Double(H(n-1)));
}

Draw(H(n));
}
hilbert.curve(n=3)

```


## Part 4: The Koch Snowflake

```{r}
#Script by A. Roberts, 2013.
#To run the script, copy and paste it onto the R command line, and press ,<enter>.

KochSnowflakeExample <- function(){
iterate <- function(T,i){
A = T[ ,1]; B=T[ ,2]; C = T[,3];
if (i == 1){
d = (A + B)/2; h = (C-d); d = d-(1/3)*h;
e = (2/3)*B + (1/3)*A; f = (1/3)*B + (2/3)*A;
}

if (i == 2){
d = B; e = (2/3)*B + (1/3)*C; f = (2/3)*B + (1/3)*A;
}

if (i == 3){
d = (B + C)/2; h = (A-d); d = d-(1/3)*h;
e = (2/3)*C + (1/3)*B; f = (1/3)*C + (2/3)*B;
}

if (i == 4){
d = C; e = (2/3)*C + (1/3)*A; f = (2/3)*C + (1/3)*B;
}
if (i == 5){
d = (A + C)/2; h = (B-d); d = d-(1/3)*h;
e = (2/3)*A + (1/3)*C; f = (1/3)*A + (2/3)*C;
}

if (i == 6){
d = A; e = (2/3)*A + (1/3)*C; f = (2/3)*A + (1/3)*B;
}

if (i == 0){
d = A; e = B; f = C;
}

Tnew = cbind(d,e,f)
return(Tnew); #Return a smaller triangle.
}

draw <- function(T, col=rgb(0,0,0),border=rgb(0,0,0)){
polygon(T[1,],T[2,],col=col,border=border)
}

Iterate = function(T,v,col=rgb(0,0,0),border=rgb(0,0,0)){
for (i in v) T = iterate(T,i);
draw(T,col=col,border=border);
}

#The vertices of the initial triangle:
A = matrix(c(1,0),2,1);
B = matrix(c(cos(2*pi/3), sin(2*pi/3)),2,1);
C = matrix(c(cos(2*pi/3),-sin(2*pi/3)),2,1);
T0 = cbind(A,B,C);

plot(numeric(0),xlim=c(-1.1,1.1),ylim=c(-1.1,1.1),axes=FALSE,frame=FALSE,ann=FALSE);
par(mar=c(0,0,0,0),bg=rgb(1,1,1));
par(usr=c(-1.1,1.1,-1.1,1.1));

# Draw snowflake:
for (i in 0:6) for (j in 0:6) for (k in 0:6) for (l in 0:6) Iterate(T0,c(i,j,k,l));
}

# Run the example
KochSnowflakeExample()


```

## Part 5: Sierpinski Carpet with ggplot2

Need to investigate how grid works (e.g. how to display the plots on one page).
The code uses quartz() which is (apparently) a MacOS specific display driver.
I added the first print(p) to get any displaay.  Eventually I gave up trying to
make this code work and added in the gridExtra package to make the plots. The
original logic would make the plots but it would plot a new page for each of the
four viewport specifications.

```{r}
#Written by Allan Roberts, Feb 2014
library(ggplot2)
library(grid)
library(gridExtra)
SierpinskiCarpet <- function(k){
Iterate <- function(M){
A <- cbind(M,M,M);
B <- cbind(M,0*M,M);
return(rbind(A,B,A))
}
M <- as.matrix(1)
for (i in 1:k) M <- Iterate(M);
n <- dim(M)[1]
X <- numeric(n)
Y <- numeric(n)
I <- numeric(n)
for (i in 1:n) for (j in 1:n){
X[i + (j-1)*n] <- i;
Y[i + (j-1)*n] <- j;
I[i + (j-1)*n] <- M[i,j];
}
DATA <- data.frame(X,Y,I)
p <- ggplot(DATA,aes(x=X,y=Y,fill=I))
### print(p)
p <- p + geom_tile() + theme_bw() + scale_fill_gradient(high=rgb(0,0,0),low=rgb(1,1,1))
p <-p+ theme(legend.position=0) + theme(panel.grid = element_blank())
p <- p+ theme(axis.text = element_blank()) + theme(axis.ticks = element_blank())
p <- p+ theme(axis.title = element_blank()) + theme(panel.border = element_blank());
return(p)
}
### A <- viewport(0.25,0.75,0.45,0.45)
### B <- viewport(0.75,0.75,0.45,0.45)
### C <- viewport(0.25,0.25,0.45,0.45)
### D <- viewport(0.75,0.25,0.45,0.45)
### quartz(height=6,width=6)
### print(SierpinskiCarpet(1),vp=A)
### print(SierpinskiCarpet(2),vp=B)
### print(SierpinskiCarpet(3),vp=C)
### print(SierpinskiCarpet(4),vp=D)
grid.arrange(SierpinskiCarpet(1),
             SierpinskiCarpet(2),
             SierpinskiCarpet(3),
             SierpinskiCarpet(4),
             nrow=2)

```

All done
