---
title: "Koch Snowflake"
author: "R. G. Cronce"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Koch Snowflake

A fractal shape with infinite perimeter and finite area.  Generated by removing the center
third of a line and replacing with the upper part of equilateral triangle.  Iteratively
process each of the two legs of the triangle.  When done with hexagonal symmetry this yields
something of a fractal snowflake.

[Website](https://bmscblog.wordpress.com/2013/10/04/fractals-with-r-part-4-the-koch-snowflake/)

code to generate the snowflake - directly copied to here.

```{r}
# A short article and
# Script by A. Roberts, 2013.
# To run the script, copy and paste it onto the R command line, and press ,<enter>.

KochSnowflakeExample <- function(){
iterate <- function(T,i){
A = T[ ,1]; B=T[ ,2]; C = T[,3];
if (i == 1){
d = (A + B)/2; h = (C-d); d = d-(1/3)*h;
e = (2/3)*B + (1/3)*A; f = (1/3)*B + (2/3)*A;
}

if (i == 2){
d = B; e = (2/3)*B + (1/3)*C; f = (2/3)*B + (1/3)*A;
}

if (i == 3){
d = (B + C)/2; h = (A-d); d = d-(1/3)*h;
e = (2/3)*C + (1/3)*B; f = (1/3)*C + (2/3)*B;
}

if (i == 4){
d = C; e = (2/3)*C + (1/3)*A; f = (2/3)*C + (1/3)*B;
}

if (i == 5){
d = (A + C)/2; h = (B-d); d = d-(1/3)*h;
e = (2/3)*A + (1/3)*C; f = (1/3)*A + (2/3)*C;
}

if (i == 6){
d = A; e = (2/3)*A + (1/3)*C; f = (2/3)*A + (1/3)*B;
}

if (i == 0){
d = A; e = B; f = C;
}

Tnew = cbind(d,e,f)
return(Tnew); #Return a smaller triangle.
}

draw <- function(T, col=rgb(0,0,0),border=rgb(0,0,0)){
polygon(T[1,],T[2,],col=col,border=border)
}

Iterate = function(T,v,col=rgb(0,0,0),border=rgb(0,0,0)){
for (i in v) T = iterate(T,i);
draw(T,col=col,border=border);
}

#The vertices of the initial triangle:
A = matrix(c(1,0),2,1);
B = matrix(c(cos(2*pi/3), sin(2*pi/3)),2,1);
C = matrix(c(cos(2*pi/3),-sin(2*pi/3)),2,1);
T0 = cbind(A,B,C);

plot(numeric(0),xlim=c(-1.1,1.1),ylim=c(-1.1,1.1),axes=FALSE,frame=FALSE,ann=FALSE);
# RGC: needed to modify left/right margins to make aspect ratio correct (e.g. square)
# or change the scaling.
# Using the margins:
#    For HTML output a margin of 4 is about right
#    For RStudio output a margin of 7 is about right
# Using the scaling, xlimits of +/- 1.6 seem about right all the time
par(mar=c(0,0,0,0),bg=rgb(1,1,1));
par(usr=c(-1.6,1.6,-1.1,1.1));

#Draw snowflake:
for (i in 0:6) for (j in 0:6) for (k in 0:6) for (l in 0:6) Iterate(T0,c(i,j,k,l));
}

# Run the example.
KochSnowflakeExample()

```

and there it is.