---
title: "Logistic Map"
author: "R. G. Cronce"
date: "10/2/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Logistic Map

Aka the predator/prey equation - one of the simplest chaotic systems that occurs in
virtually all of the articles about chaos.  Representatitve of a system with competing
birth/death rate (r) and a population (x).  The iteration is:

    x[n+1] = r*x[n]*(1-x[n])

Where:

    x[n] = population at time n, in the range 0<x<1
    x[n+1] = population at time n+1
    r    = reproduction rate

When r is small the population dies out.  When r has a medium value `(1<r<3)` there is
a single steady state.  For r>3 there is a bifurcation in the population level in subsequent
generations. As r increases the bifurcations exhibit period doubling and eventual chaos.  For
small regions the response returns to simple bifurcation and then chaos resumes above
that level of r.  In order to see the chaotic behaviour you need to explore the same value
of r for several iterations before moving on to the next value of r.


```{r}
logistic = function(x,r) {
    x=r*x*(1-x)
}

# Range of r values for exploration
r0=2.5
r1=4.0
# Number of repeats at each value of r
nrepeat=200

# Random starting point
z=runif(1)
# Iterate to get to a stationary point (if it exists) or on the attractor
for (i in 1:nrepeat) { z=logistic(z,r0) }
# Range of r values
r=seq(r0,r1,length=1000)
# Use nrepeat points at each value of r
x=rep(r,each=nrepeat)
y=0*x
for (i in seq_along(x)) {
    z=logistic(z,x[i])
    y[i]=z
}
plot(x,y,pch=19,cex=0.1,xlab="Growth Rate",ylab="Population",main="Logistic Map Chaos")
polygon(c(3.4,3.65,3.65,3.4),c(0.29,0.29,0.91,0.91),border="red",lwd=3)


```

This is the typical logistic map bifurcation diagram.

## Self-similarity

The trajectory bifurcates for r>3 and then again for r just above 3.4.  This
period doubling and self-similarity can be seen below -- this is the same
region as shown above inside the red box.

```{r}
# Range of r values for exploration
r0=3.4
r1=3.65
# Number of repeats at each value of r
nrepeat=200

# Random starting point
z=runif(1)
# Iterate to get to a stationary point (if it exists) or on the attractor
for (i in 1:nrepeat) { z=logistic(z,r0) }
# Range of r values
r=seq(r0,r1,length=1000)
# Use nrepeat points at each value of r
x=rep(r,each=nrepeat)
y=0*x
for (i in seq_along(x)) {
    z=logistic(z,x[i])
    y[i]=z
}
plot(x,y,pch=19,cex=0.1,xlab="Growth Rate",ylab="Population",main="Logistic Map Chaos")
polygon(c(3.63,3.636,3.636,3.63),c(0.82,0.82,0.85,0.85),border="red",lwd=3)


```

## Stability within Chaos

The small area boxed in red above is examined below.  Note the self-similarity to the
previous plots and the abrupt transition from chaos to/from well-ordered behavior.

```{r}
# Range of r values for exploration
r0=3.63
r1=3.636
# Number of repeats at each value of r
nrepeat=600

# Random starting point
z=runif(1)
# Iterate to get to a stationary point (if it exists) or on the attractor
for (i in 1:nrepeat) { z=logistic(z,r0) }
# Range of r values
r=seq(r0,r1,length=1000)
# Use nrepeat points at each value of r
x=rep(r,each=nrepeat)
y=0*x
for (i in seq_along(x)) {
    z=logistic(z,x[i])
    y[i]=z
}
plot(x,y,pch=19,cex=0.1,
     ylim=c(0.82,0.85),
     xlab="Growth Rate",ylab="Population",main="Logistic Map Chaos")

```
