---
title: "Sierpinski Gasket"
author: "R. G. Cronce"
date: "10/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Sierpinski Gasket

Aka the Sierpinski triangle - one of the simplest fractal shapes that can be generated.
Start with an equilateral triangle and remove the center formed by connecting the midpoints
of each of the sides.  The result is three smaller equilateral triangles surrounding an
upside down triangle.  Repeat the process for each of the three outer triangles and recurse.

An alternate implementation is based on a stochastic process.  Fix the three vertices of
the generating triangle and a random starting point.  At each step, select one of the
three vertices at random and move half-way to the corresponding vertex.  Plot a point
and repeat ad nauseum.  As the number of points is increased the gasket will "magically"
appear.

```{r}
# Create vertices for equilateral triangle with unit sides
x=c(0,0.5,1)
y=c(0,sqrt(3)/2,0)

# Make the response repeatable
set.seed(42)

# Generate a random starting point
xval=runif(1)
yval=runif(1)

# Iterate to create a large number of points
xxx=c()
yyy=c()
for (i in 1:10000) {
d=sample(3,1)
xval=(xval+x[d])/2
yval=(yval+y[d])/2
xxx=c(xxx,xval)
yyy=c(yyy,yval)
}

# Plot the results (suppress display of the axes)
par(mar=c(0,0,1,0))
plot(xxx,yyy,pch=15,cex=0.5,asp=1,main='Sierpinski Gasket')

```

This is the typical Sierpinski Gasket (Triangle).
